On the effects of partying on next-day sports performance

### 2017-06-11 writing/draft.tex
Added heat map for last-game location vs. next day opponent.

\section{Defining a Measure for Partying}
One of the hardest parts of this analysis is that partying is a latent variable.
In order to determine when players partied, one may consider looking at social media data for direct mentions of partying, either by athletes or witnesses. However, due to the intense public scrutiny that athletes face, there is a strong reluctance to disclose their personal lives on social media. There are notable exceptions, such
as Oddel Beckham Jr.'s recent trip to Miami.\citep{bleler}
However, these reports are few and far enough between that picking up statistical evidence would be fruitless.
We develop two approaches: a discrete indicator and a continuous measure of nightlife.

@@

\begin{figure}
%\includegraphics[scale=0.5]{travel_dist_by_party}
  \centering
  \includegraphics[scale=0.55]{travel_dist_density_by_party}
  \label{fig: density by party}
  \caption{We examine travel distance to current game location as a function of
    whether the last game was in a (non) party city. We see that in general,
  travel distance follows a similar distribution regardless of last game location, i.e. 
  it's not as though players encounter significantly
  more travel fatigue after playing in a party city.}
\end{figure}

@@

\begin{sidewaysfigure}
  \hspace{-55pt}
  \includegraphics[scale=0.7]{Party_by_Team_Location}
\end{sidewaysfigure}

@@

>>>>

\section{Defining a Measure for Partying}
A key assumption in our analysis is that next-day opponent is reasonably uncorrelated with
last game location. We plot a heat-map displaying the relationship between last game location
and next game location, and further break this down according to whether the game(s)
were played at home or away.

\begin{figure}
  \hspace{-44pt}
  \includegraphics[scale=0.45]{next_day_opponent}
  \caption{We examine the relationship between last game location and next day opponent in the NBA. Starting from the top left, notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home. This makes sense; 1/2 of all games are played at home, and 1/2 are played away, so clearly each team is most likely to be found playing a home game for any given game. When exactly one of the last game or current game was played away,
we observe that some combinations cease to exist in our data; e.g. we never observe Charlotte Bobcats play Utah Jazz on the road and then fly directly back home for a game. Lastly, notice the patterns exhibited when last game is played on the road, and current game also played on the road. We see that when last game location was in Los Angeles, next day opponent likely to be a west coast team. There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because players ``gain'' time when flying east-to-west.} \end{figure}

Our takeaway from the figure above is that indeed there is a correlation between last game location
and next day opponent, in particular this is true for west coast teams. We hypothesize that
this is due to practical limitations on flight schedules. The fact that there is a correlation between last game location and next day opponent increases the motivation for conditioning on spread, which accounts may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.

\paragraph{Partying as a latent variable}
One of the hardest parts of this analysis is that partying is a latent variable.
In order to determine when players partied, one may consider looking at social media data for direct mentions of partying, either by athletes or witnesses. However, due to the intense public scrutiny that athletes face, there is a strong reluctance to disclose their personal lives on social media. There are notable exceptions, such
as Oddel Beckham Jr.'s recent trip to Miami.\citep{bleler}
However, these reports are few and far enough between that picking up statistical evidence would be fruitless.
We develop two approaches: a discrete indicator and a continuous measure of nightlife.

@@

\begin{figure}
%\includegraphics[scale=0.5]{travel_dist_by_party}
  \centering
  \includegraphics[scale=0.55]{travel_dist_density_by_party}
  \label{fig: density by party}
  \caption{Within the NBA, we examine travel distance to current game location as a function of
    whether the last game was in a (non) party city. We see that in general,
  travel distance follows a similar distribution regardless of last game location, i.e. 
  it's not as though players encounter significantly
  more travel fatigue after playing in a party city. We mention that the same plot is not interesting to look at within the MLB, since games are played in series, and therefore most travel-distances between games are zero miles.}
\end{figure}

@@

\begin{sidewaysfigure}
  \hspace{-55pt}
  \includegraphics[scale=0.7,angle=180]{Party_by_Team_Location}
\end{sidewaysfigure}

@@

\section{Appendix}

\begin{figure}   
  \centering
  \hspace{-55pt}
  \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
  \caption{Continuous measure of nightlife by team location for MLB. What's interesting here is that because New York Yankees are in the Bronx, they show up as having few musicians and sound recording studios. We might wonder if this is a bad thing, since with NBA New York city was deemed a major party city. It turns out that when we define a discrete indicator of party based on New York or Los Angeles, it is not at all significant, and in fact the sign is incorrect. When we use our continuous
 measure of nightlife, its significant, and it helps inform our profitable betting model.
This is good news for our proxy, although begs the question why New York should be treated so differently by MLB players when compared with NBA players. One point worth noting is that we haven't yet made the important distinction that in the MLB, the Los Angeles Angels are in Orange County, for example, which is much less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los Angeles.}
\end{figure}

\begin{figure}   
  \centering
  \hspace{-55pt}
  \includegraphics[scale=0.5]{next_day_opponent_mlb} 
\end{figure}
